### 2025-05-16 chapters/mpif/ms.Rnw
Addressing pre-defense comment: "Chapter titles must be in title case"
Fixing pre-defense review: "All figure and table captions must match word for word up to the first punctuation mark."

\chapter{Iterating marginalized Bayes maps for likelihood maximization}
\label{chpt:mpif}

@@

ggplot(df_all_mu) +
  geom_point(
    aes(x = mu1, y = mu2, col = as.factor(step)),
    pch = 18, size = 1
  ) +
  geom_path(
    data = df_all_Sigma,
    aes(x = X1, y = X2, col = as.factor(step))
  ) +
  geom_path(
    data = like_elipse,
    aes(x = X1, y = X2), linetype = 'dashed'
  ) +
  geom_point(x = 0, y = 0, col = 'red', pch = 'x') +
  theme_bw() +
  scale_x_continuous(limits = c(-3.12, 13.5)) +
  scale_y_continuous(limits = c(-3.12, 13.5)) +
  theme(aspect.ratio = 1) +
  labs(color = 'Step') +
  scale_color_manual(
    values = c(
      "#800026",
      "#bd0026",
      "#e31a1c",
      "#fc4e2a",
      "#fd8d3c",
      "#feb24c"
    )
  ) +
  labs(x = TeX('$\\phi$'), y = TeX('$\\psi$')) +
  theme(legend.position = 'right') +
  facet_wrap(~group, labeller = labeller(group = method_names)) + 
  theme(plot.margin = unit(c(0, 0, 0, 0), 'cm'))
@
\caption[Data cloning compared to marginalized data cloning for Gaussian densities.]{\label{fig:DC}Data cloning and Marginalized data cloning for two parameter model with Gaussian likelihoods and priors. The ellipses show the region of the parameter distribution that contains $95\%$ of the probability mass of the distribution. The black dashed line shows this region for the likelihood surface, and the red ``x" marks the MLE.}
\end{figure}

@@

cowplot::plot_grid(depletion_1, depletion_2, nrow = 2, labels = "AUTO", rel_heights = c(0.425, 0.575))
@
\caption[Single unit iteration of MPIF on toy model.]{\label{fig:depletion}Updating parameter distributions with a single $u = 1$ iteration of both versions of Algorithm~\ref{alg:mpif}. (A) The total number of unique particles representing each parameter. The dashed horizontal line shows that MPIF maintains the number of particles for $\Psi_2$ over time. (B) Parameter particle swarm of a single update with and without marginalization compared to the true posterior distribution.}
\end{figure}

@@

cowplot::plot_grid(gg_smallU, gg_bigU, nrow = 2, rel_heights = c(0.4, 0.6))
@
\caption[MPIF compared to PIF on Gompertz population model.]{\label{fig:gomp}Comparison of the MPIF and PIF algorithms for fitting the stochastic Gompertz population model. 
The solid horizontal line shows the true maximum likelihood, determined via the Kalman filter and a numeric optimizer, an intractable approach for high-dimensional parameter spaces $(U > 100)$, so in these cases the dashed line indicates the likelihood at the data-generating parameters. 
Each algorithm used 50 unique starting points.
Vertical bars span the tenth percentile to the maximum likelihood values; lowest values were excluded due to poor likelihoods from the PIF algorithm.}
\end{figure}

@@

print(gg)
@
\caption[MPIF comparison using mechanistic model for UK measles incidence data.]{\label{fig:measles1} Log-likelihoods yielded by the PIF and MPIF algorithms for the mechanistic measles model to the UK data. 
Rows correspond to a different number of particles $J$ used in Algorithm~\ref{alg:mpif}. 
The log-likelihood is evaluated at iterations 20, 56, 92, 128, 164, and 200, marked by vertical grey lines.}
\end{figure}

>>>>

\chapter{Iterating Marginalized Bayes Maps for Likelihood Maximization}
\label{chpt:mpif}

@@

ggplot(df_all_mu) +
  geom_point(
    aes(x = mu1, y = mu2, col = as.factor(step)),
    pch = 18, size = 1
  ) +
  geom_path(
    data = df_all_Sigma,
    aes(x = X1, y = X2, col = as.factor(step))
  ) +
  geom_path(
    data = like_elipse,
    aes(x = X1, y = X2), linetype = 'dashed'
  ) +
  geom_point(x = 0, y = 0, col = 'red', pch = 'x') +
  theme_bw() +
  scale_x_continuous(limits = c(-3.12, 13.5)) +
  scale_y_continuous(limits = c(-3.12, 13.5)) +
  theme(aspect.ratio = 1) +
  labs(color = 'Step') +
  scale_color_manual(
    values = c(
      "#800026",
      "#bd0026",
      "#e31a1c",
      "#fc4e2a",
      "#fd8d3c",
      "#feb24c"
    )
  ) +
  labs(x = TeX('$\\phi$'), y = TeX('$\\psi$')) +
  theme(legend.position = 'right') +
  facet_wrap(~group, labeller = labeller(group = method_names)) + 
  theme(plot.margin = unit(c(0, 0, 0, 0), 'cm'))
@
\caption[Data cloning compared to marginalized data cloning for Gaussian densities.]{\label{fig:DC}Data cloning compared to marginalized data cloning for Gaussian densities. The ellipses show the region of the parameter distribution that contains $95\%$ of the probability mass of the distribution. The black dashed line shows this region for the likelihood surface, and the red ``x" marks the MLE.}
\end{figure}

@@

cowplot::plot_grid(depletion_1, depletion_2, nrow = 2, labels = "AUTO", rel_heights = c(0.425, 0.575))
@
\caption[Updating parameter distributions with a single $u = 1$ iteration of both versions of Algorithm~\ref{alg:mpif}.]{\label{fig:depletion}Updating parameter distributions with a single $u = 1$ iteration of both versions of Algorithm~\ref{alg:mpif}. (A) The total number of unique particles representing each parameter. The dashed horizontal line shows that MPIF maintains the number of particles for $\Psi_2$ over time. (B) Parameter particle swarm of a single update with and without marginalization compared to the true posterior distribution.}
\end{figure}

@@

cowplot::plot_grid(gg_smallU, gg_bigU, nrow = 2, rel_heights = c(0.4, 0.6))
@
\caption[Comparing the MPIF and PIF algorithms for fitting the stochastic Gompertz model.]{\label{fig:gomp}Comparing the MPIF and PIF algorithms for fitting the stochastic Gompertz model.
The solid horizontal line shows the true maximum likelihood, determined via the Kalman filter and a numeric optimizer, an intractable approach for high-dimensional parameter spaces $(U > 100)$, so in these cases the dashed line indicates the likelihood at the data-generating parameters. 
Each algorithm used 50 unique starting points.
Vertical bars span the tenth percentile to the maximum likelihood values; lowest values were excluded due to poor likelihoods from the PIF algorithm.}
\end{figure}

@@

print(gg)
@
\caption[Comparing the MPIF and PIF algorithms using a mechanistic model for measles.]{\label{fig:measles1} 
Comparing the MPIF and PIF algorithms using a mechanistic model for measles.
Rows correspond to a different number of particles $J$ used in Algorithm~\ref{alg:mpif}. 
The log-likelihood is evaluated at iterations 20, 56, 92, 128, 164, and 200, marked by vertical grey lines.}
\end{figure}
